The state of knowledge for marine invertebrate biodiversity in the continental US

### 2022-02-04 manuscript.Rmd
add bibliography

---
title: "The state of knowledge for marine invertebrate biodiversity in the continental US"
date: ' ' #'`r format(Sys.time())`'
output:
  bookdown::pdf_document2:
    includes:
    #  before_body: ../template/doc_prefix.tex
      in_header: preamble.tex
    keep_tex: yes
    latex_engine: xelatex
    number_sections: no
    toc: no
fontsize: 12pt
link-citations: yes
csl: nature.csl
bibliography: /home/francois/Library/dark-plankton.bib
abstract: Museum collections have been recording our knowledge about the diversity and distribution of species for about 250 years. Until recently, this information was fragmented across institutions and difficult to access or synthesize. The digitization of biodiversity information is enabling the synthesis of collection data. Here we use publicly accessible digital biodiversity data to asses how well we know the marine invertebrate fauna of the continental United States. Globally, marine inverterbrates comprise about 75% of described marine species. We document `r format(n_distinct(combined_species$worms_valid_name), big.mark = ",")` species from the continental US waters in digital databases (OBIS and iDigBio), and explore spatial and temporal patterns of their distribution and collection. Databases do well at the fauna level, capturing ~xxx% of the species diversity. However, sampling bias obscure broad diversity patterns, such as differences between coasts and across latitude. Our knowledge about species richness and geographic range is limited by sampling effort, with `r prop_10_specimens_species`% of species known from $\leq$ 10 samples. While recent developments in high-throughput sequencing promise to greatly speed up documentation of marine biodiversity, interpretation of sequences requires anchoring results anchoring results to identified vouchers. DNA barcode sequences (COI) are presently available for only `r format_output(prop_spp_barcoded$p_barcoded * 100)`% of the US marine invertebrates, and many of these are from samples taken outside the US. Over a third of the barcoded species are complexes of of multiple cryptic lineages, indicating that the fauna is substantially richer than the above number indicate and needs to be taxonomically reevaluated as a whole. Substantially increased sampling, coordinated at the national scale, is needed to provide a meaningful baseline for marine biodiversity and for metabarcoding and metagenomics studies. This data will allow the detection of the impacts of climate change and other anthropogenic strains on marine biodiversity. Advances in computing, information systems, and sequencing technologies, have set the stage for a broad understanding of the dynamics of marine biodiversity, but this potentially is limited by the availablility of current samples, digitization, and taxonomic effort.
---

>>>>

---
title: "The state of knowledge for marine invertebrate biodiversity in the continental US"
date: ' ' #'`r format(Sys.time())`'
output:
  bookdown::pdf_document2:
    includes:
    #  before_body: ../template/doc_prefix.tex
      in_header: preamble.tex
    keep_tex: yes
    latex_engine: xelatex
    number_sections: no
    toc: no
fontsize: 12pt
link-citations: yes
csl: nature.csl
bibliography: sok.bib
abstract: Museum collections have been recording our knowledge about the diversity and distribution of species for about 250 years. Until recently, this information was fragmented across institutions and difficult to access or synthesize. The digitization of biodiversity information is enabling the synthesis of collection data. Here we use publicly accessible digital biodiversity data to asses how well we know the marine invertebrate fauna of the continental United States. Globally, marine inverterbrates comprise about 75% of described marine species. We document `r format(n_distinct(combined_species$worms_valid_name), big.mark = ",")` species from the continental US waters in digital databases (OBIS and iDigBio), and explore spatial and temporal patterns of their distribution and collection. Databases do well at the fauna level, capturing ~xxx% of the species diversity. However, sampling bias obscure broad diversity patterns, such as differences between coasts and across latitude. Our knowledge about species richness and geographic range is limited by sampling effort, with `r prop_10_specimens_species`% of species known from $\leq$ 10 samples. While recent developments in high-throughput sequencing promise to greatly speed up documentation of marine biodiversity, interpretation of sequences requires anchoring results anchoring results to identified vouchers. DNA barcode sequences (COI) are presently available for only `r format_output(prop_spp_barcoded$p_barcoded * 100)`% of the US marine invertebrates, and many of these are from samples taken outside the US. Over a third of the barcoded species are complexes of of multiple cryptic lineages, indicating that the fauna is substantially richer than the above number indicate and needs to be taxonomically reevaluated as a whole. Substantially increased sampling, coordinated at the national scale, is needed to provide a meaningful baseline for marine biodiversity and for metabarcoding and metagenomics studies. This data will allow the detection of the impacts of climate change and other anthropogenic strains on marine biodiversity. Advances in computing, information systems, and sequencing technologies, have set the stage for a broad understanding of the dynamics of marine biodiversity, but this potentially is limited by the availablility of current samples, digitization, and taxonomic effort.
---
